\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-08-07 coq-example.tex
Change the final example illustration parts.

 Inherits subst.
 ...
 Extend FInductive has_type 
  : context -> term -> type := ...
 Extend FTheorem subst_lemma.
  ... 
 Qed.
(* ... and more, end with type safety *)
(* Next, Invoking Mixin Command *)
EndFamily.
Family STLC_bool_prod 
    extends STLC 
    using STLC_bool using STLC_prod.
\end{lstlisting}
\includegraphics[width=0.5\textwidth]{coqexmaple/STLC-inheritances.pdf}
\caption{Example STLC extended with product, and Mixin with \cref{fig:STLC-example2}}
\end{wrapfigure}

@@

Our emphasis here is that, we singled out some language features from
STLC and prove the type safety for each feature separately, following
the style of the examples in MTC~\cite{delaware2013,forsta2020}.
And we use a \textit{not-yet-formally-defined} \textbf{mixin} feature
to mix the semantics and properties of two language features---product
and boolean (in \cref{fig:STLC-example})---with vanilla STLC.

@@

Our formulation for bare STLC is around 400+ LOCs; the two families implementing product and boolean features both take around 250+ LOCs. 
% need comparison with the original implementation

@@

\subsection{Abstract Interpreters for \texttt{Imp}}\YZ{Try to have two runnable interpreter -- choose another simple one like constant propagation. and remove LangMore}
The second example is adapted and modified from \citet{zm2017}.
Contrary to our first example, here, we use a big-step interpreter with fuel
as the operational semantics of an imperative language with
side effects, and we specify the abstract interpretation and prove its
soundness, with some postulation on both computation and property.
Then we extend the base language with new features, and we instantiate
the postulation on computation for
abstract interpreters.
Thanks to the compilation to Coq module, we can directly run the resulting
abstract interpreters.

\begin{figure}[!htb]
  \includegraphics[width=\columnwidth]{coqexmaple/Family-Lang-Imp2.pdf}
  \caption{Inheritance Diagram for Abstract Interpretation Example. Black arrows are inheritance. Purple arrow are extending. Green arrows are overriding. Grey box are sealed families.}\label{fig:abstract-interpretation-example}
\end{figure}

We construct four families, in one inheritance chain, illustrated in \cref{fig:abstract-interpretation-example}.\YZ{Why is it useful to define these concrete and abstract interpreters as an inheritance chain?}\EDJreply{I don't think these examples are here to show the example themselves are useful. They are here to show family inheritance is useful, and can extend stuff to all sorts of things. I also add some details below on how each family can be extended and what each family means.}\YZreply{Why is it desired to define the abstract interpreter as an extension of the concrete interpreter family? What is lost by putting the two interpreters in the same family?}\EDJreply{Nothing is lost. This is just simulating people starting with bare (big-step) operational semantic. Because they may not working on abstract interpretation from the first place. Those people may extend "Lang" with other stuff like another type-safety proof or something.}

The first family "Lang" define the big step operator (using fuel) of a
while language, with expression and stmt defined upon expression. Final "Result" can be error, "out-of-fuel" or a legitmate "State". It takes about 200+ LOCs. 
% We
% can
\YZ{In general, in a result section like this, you want to show what
you *have done* rather than what you *can do*.}\EDJreply{Ok. deleted. I initially want to emphasize ``I **can** use arbitrary implementation for this environment mapping interface'' but I don't know how to say that without using ``can''.}
% swap the implementation of "State" to have different memory/state
% management by extending family "Lang".
\YZ{What is 'different
memory/state management'?}\EDJreply{Actually I should say environment mapping/dictionary. I instantiate it with function(closure) as environment mapping in "LangRun". It is possible to use other mapping like Coq's primitive dictionary.}

@@

The third Family "LangMore" extend "LangwAbs" with nat constant and
addition, and if-then-else control flow, of course retaining all the
soundness theorem. It takes about 200+ LOCs. This is one example of
using family inheritance to support new language feature, and compatible
with the existent reasoning in "LangwAbs.AbsInterpreter".

The fourth family "LangRun" instantiate some of the postulates of the
"LangMore" and recover computational information of "analyze". The "absStateData" is overridden by a concrete definition. The resulting abstract interpreter is expected
to act as a type-analyzer. It takes about 200+ LOCs. This is one example
of instantiate the detail computation of abstract interpretation.

And at the end, we compile LangRun into module and use Coq to run it on
some simple queries. This "LangRun" also illustrates that we still have
computability in the presence of family.\YZ{Can there be a diagram (a la Figure 1 in Familia paper) illustrating the relationships between (and functionalities of) the different families?}\EDJreply{Done. Please check}

>>>>

 Inherits subst.
 ...
 Extend FInductive has_type 
  : context -> term -> type := ...
 Extend FTheorem subst_lemma.
  ... 
 Qed.
(* ... and more, end with type safety *)
(* Next, Invoking Mixin Command *)
EndFamily.
Family STLC_bool_prod 
    extends STLC 
    using STLC_bool using STLC_prod.
Family STLC_bool_prod_sum 
    extends STLC 
    using STLC_sum using STLC_bool_prod.
Family STLC_bool_prod_sum_fix 
    extends STLC 
    using STLC_bool_prod_sum using STLC_fix.
(* ... and more, 
   there are 11 mixins families ... *)
\end{lstlisting}
\includegraphics[width=0.5\textwidth]{coqexmaple/STLC-inheritances.pdf}
\caption{Example STLC extended with product, and Mixin with \cref{fig:STLC-example2}}
\end{wrapfigure}

@@

Our emphasis here is that, we singled out some language features from
STLC and prove the type safety for each feature separately, following
the style of the examples in MTC~\cite{delaware2013,forsta2020}.
And we use a \textit{not-yet-formally-defined} \textbf{mixin} feature
to mix the semantics and properties of language features---product
, boolean (in \cref{fig:STLC-example}), sum type and iso-recursive type---with vanilla STLC.

@@

Our formulation for bare STLC is around 400+ LOCs; each families implementing product and boolean features both take around 300+ LOCs. 
% need comparison with the original implementation

@@

\subsection{Abstract Interpreters for \texttt{Imp}}\YZ{Try to have two runnable interpreter -- choose another simple one like constant propagation. and remove LangMore}
The second example is adapted and modified from \citet{zm2017}.
Contrary to our first example, here, we use a big-step interpreter with fuel
as the operational semantics of an imperative language with
side effects, and we specify the abstract interpretation and prove its
soundness, with some postulation on both computation and property. Then we instantiate our abstract interpretation with two concrete computation---one for type inference and the other for constant propagation. Our analyzer will return the analysis result about the end of the program, e.g. our constant propagator will return what constant value each variable is at the end of the program. Thanks to the compilation to Coq module, we can directly run the resulting
abstract interpreters.

% Then we extend the base language with new features, and we instantiate
% the postulation on computation for
% abstract interpreters.

\begin{figure}[!htb]
  \includegraphics[width=\columnwidth]{coqexmaple/Family-Lang-Imp3.pdf}
  \caption{Inheritance Diagram for Abstract Interpretation Example. Black arrows are inheritance. Purple arrow are extending. Green arrows mean there are overriding.}\label{fig:abstract-interpretation-example}
\end{figure}

We construct four families, illustrated in \cref{fig:abstract-interpretation-example}.\YZ{Why is it useful to define these concrete and abstract interpreters as an inheritance chain?}\EDJreply{I don't think these examples are here to show the example themselves are useful. They are here to show family inheritance is useful, and can extend stuff to all sorts of things. I also add some details below on how each family can be extended and what each family means.}\YZreply{Why is it desired to define the abstract interpreter as an extension of the concrete interpreter family? What is lost by putting the two interpreters in the same family?}\EDJreply{Nothing is lost. This is just simulating people starting with bare (big-step) operational semantic. Because they may not working on abstract interpretation from the first place. Those people may extend "Lang" with other stuff like another type-safety proof or something.}
The first family "Lang" define the big step operator (using fuel) of a
while language, with expression and stmt defined upon expression. Final "Result" can be error, "out-of-fuel" or a legitmate "State". It takes about 200 LOCs. 
% We
% can
\YZ{In general, in a result section like this, you want to show what
you *have done* rather than what you *can do*.}\EDJreply{Ok. deleted. I initially want to emphasize ``I **can** use arbitrary implementation for this environment mapping interface'' but I don't know how to say that without using ``can''.}
% swap the implementation of "State" to have different memory/state
% management by extending family "Lang".
\YZ{What is 'different
memory/state management'?}\EDJreply{Actually I should say environment mapping/dictionary. I instantiate it with function(closure) as environment mapping in "LangRun". It is possible to use other mapping like Coq's primitive dictionary.}

@@

% The third Family "LangMore" extend "LangwAbs" with nat constant and
% addition, and if-then-else control flow, of course retaining all the
% soundness theorem. It takes about 200+ LOCs. This is one example of
% using family inheritance to support new language feature, and compatible
% with the existent reasoning in "LangwAbs.AbsInterpreter".

The third family and the fourth family "LangTyping" and "LangCP" instantiate some of the postulates of the
"LangwAbs" and recover computational information of "analyze". The corresponding fields in "absStateData" is overridden by a concrete definition. The resulting abstract interpreter is expected
to act as a type-analyzer and constant-propagator respectively. It takes about 200+ LOCs. 
% These are two examples of instantiate the detail computation of abstract interpretation.

And at the end, the compiled module for "LangTyping" and "LangCP" is put into tests over simple queries. These "LangTyping" and "LangCP" also illustrates that we still have
computability in the presence of family.\YZ{Can there be a diagram (a la Figure 1 in Familia paper) illustrating the relationships between (and functionalities of) the different families?}\EDJreply{Done. Please check}
